Faithful Variable Screening for High-Dimensional Convex Regression

### 2014-09-12 tex/scam/sparsistency.tex
minor edits

We construct an additive convex solution $\{\hat{d}_k\}_{k=1,...,p}$
that is zero for $k \in S^c$, where $S$ is the set of relevant
variables, and show that it satisfies the KKT
conditions for optimality of optimization~\eqref{opt:alternate_opt}. We
define $\hat{d}_k$ for $k \in S$ to be a solution to the restricted
regression (defined below). We also show that $\hat{c}_k =
0$ satisfies the optimality condition of
optimization~\eqref{opt:alternate_opt_concave} for all $k \in S^c$.

\begin{definition}
\label{def:restricted_regression}
We define the \emph{restricted regression} problem 
\[
\min_{d_k} \frac{1}{n} \Big\| Y - \sum_{k \in S} \bar{\Delta}_k d_k \Big\|_2^2 + 
   \lambda_n \sum_{k \in S} \| \bar{\Delta}_k d_k \|_\infty \quad \trm{such that} \, d_{k,1}, ..., d_{k,n-1} \geq 0
\]
where we restrict the indices $k$ in
optimization \eqref{opt:alternate_opt} to lie in the set $S$ of true
relevant variables.
\end{definition}

\begin{theorem}[Deterministic setting]
\label{thm:deterministic}
Let $\{\hat{d}_k \}_{k \in S}$ be a minimizer of the restricted regression as defined above.
Let $\hat{r} \coloneqq Y - \sum_{k \in S} \bar{\Delta}_k \hat{d}_k$ be the restricted regression residual. 
Suppose for all $k\in S^c$, for all $i=1,...,n$, $\lambda_n > | \frac{1}{2n}
\hat{r}^\tran \mathbf{1}_{(i:n)}|$ where $\mathbf{1}_{(i:n)}$ is 1 on
the coordinates of the 
$i$-th largest to the $n$-th largest entries of $X_k$ and 0
elsewhere.  Then the following two statements hold.
\begin{enumerate}
\item Let $\hat{d}_k = 0$ for $k \in S^c$.  Then
  \{$\hat{d}_k\}_{k=1,...,p}$ is an optimal solution to
  optimization~\eqref{opt:alternate_opt}. Furthermore, any solution to
  the optimization program \eqref{opt:alternate_opt} must be zero on
  $S^c$.
\item For all $k \in S^c$, the solution $\hat{c}_k$ to optimization~\eqref{opt:alternate_opt_concave} must be zero.
\end{enumerate}

@@

\begin{enumerate}
\item The data $X^{(1)},..., X^{(n)} \sim F$ are iid from
a distribution $F$ that is supported and strictly positive on $\mathcal{X}=[-b,b]^p$. 
\item The response is $Y = f_0(X) + W$ where $W$ is
  independent, zero-mean noise; thus $Y^{(i)} = f_0(X^{(i)}) + W^{(i)}$.
\item The regression function $f_0$ satisfies
$f_0(X) = f_0(X_{S_0})$, where $S_0 = \{1,...,s_0\}$ is the set of
relevant variables.
\end{enumerate}

Let $\mathcal{C}^1$ denote the set of univariate convex functions
supported on $[-b,b]$, 
and let  $\mathcal{C}_1^{p}$ denote the set of convex additive functions
$\mathcal{C}_1^p \equiv \{ f \,:\, f = \sum_{k=1}^p f_k, \,
   f_k \in \mathcal{C}^1 \} $.  
Let $f^*(x) = \sum_{k=1}^p f^*_k(x_k)$ be the population risk
minimizer in $\mathcal{C}_1^p$, 
\begin{equation}
f^* \equiv \arg\min_{f \in \mathcal{C}_1^p} \E\big(f_0(X) - f^*(X)
\big)^2.
\end{equation}
Similarly, we define $\mh \mathcal{C}^1$ as the set of univariate concave functions supported on $[-b, b]$ and define
\begin{equation}
g^*_k = \arg\min_{g_k \in \mh \mathcal{C}^1} \E \big( f_0(X) - f^*(X)
- g_k(X_k) \big)^2.
\end{equation}
We let $S = \{ k = 1,...,p \,:\, f^*_k \neq 0 \trm{ or } g^*_k \neq 0\}$. By additive faithfulness (Theorem~\ref{thm:acdc_faithful}), it must be that $S_0 \subset S$.

Each of our theorems will use a subset of the following assumptions:
\begin{packed_enum}
\item[A1:] $X_S, X_{S^c}$ are independent. 
\item[A2:] $f_0$ is convex and twice-differentiable. 
\item[A3:] $\|f_0\|_\infty \leq sB$ and $\| f^*_k \|_\infty \leq B$ for all $k$.
\item[A4:] $W$ is mean-zero sub-Gaussian, independent of $X$, with sub-Gaussian scale $\sigma$, i.e. for all $t \in \R$, $\E e^{t \epsilon} \leq e^{\sigma^2 t^2 / 2}$.
\end{packed_enum}
By assumption A1, $f^*_k$ is must be zero for $k\notin \{1,...,s\}$.

We define $\alpha_f, \alpha_g$ as a measure of the signal strength of the weakest variable:
\begin{align*}
\alpha_f &= \inf_{f \in \mathcal{C}_1^p \,:\, \exists k ,\, f^*_k \neq 0 \,\wedge\, f_k = 0} 
       \Big\{ \mathbb{E} \big( f_0(X) - f(X) \big)^2 - 
        \mathbb{E} \big( f_0(X) - f^*(X) \big)^2  \Big\}\\
\alpha_g &=   \min_{k \in S \,:\, g^*_k \neq 0}
      \Big\{ \mathbb{E} \big( f_0(X) - f^*(X) \big)^2 - 
    \mathbb{E} \big( f_0(X) - f^*(X) - g^*_k(X_k) \big)^2 \Big\}
\end{align*}

Intuitively, if $\alpha_f$ is smaller, then it is easier to make a false omission in the additive convex stage of the procedure. If $\alpha_g$ is smaller, then it is easier to make a false omission in the decoupled concave stage of the procedure.

@@

\begin{theorem}[Controlling false positives]
\label{thm:false_positive}
Suppose assumptions A1-A4 hold. Define $\tilde{\sigma} \equiv \max(\sigma, B)$.

Suppose $\lambda_n \geq 8 s \tilde{\sigma}  \sqrt{ \frac{1}{n} \log^2 np}$, then with probability at least $ 1 - \frac{12}{np}$, for all $k \in S^c$, and for all $i'=1,...,n$:
\[
\lambda_n > \Big| \frac{1}{2n}\hat{r}^\tran \mathbf{1}_{(i':n)_k} \Big|
\]

And therefore，for all $k \in S^c$, both the AC solution $\hat{f}_k$, from optimization~\eqref{opt:alternate_opt}, and the DC solution $\hat{g}_k$, from optimization~\eqref{opt:alternate_opt_concave} are zero. 
\end{theorem}

The proof of Theorem~\ref{thm:false_positive} exploits independence of
$\hat{r}$ and $X_k$ from A1; when $\hat{r}$ and $X_k$ are independent, $\hat{r}^\tran \mathbf{1}_{(i':n)}$ is the sum of $n - i' +1$ random coordinates of $\hat{r}$.  We can then uses the concentration of measure result for sampling without replacement to argue that $| \frac{1}{n} \hat{r}^\tran\mathbf{1}_{(i':n)}|$ is small with high probability. The full proof of
Theorem~\ref{thm:false_positive} is in
Section~\ref{sec:false_positive_proof} of the Appendix.

>>>>

We construct an additive convex solution $\{\hat{d}_k\}_{k=1,\ldots,p}$
that is zero for $k \in S^c$, where $S$ is the set of relevant
variables, and show that it satisfies the KKT
conditions for optimality of optimization~\eqref{opt:alternate_opt}. We
define $\hat{d}_k$ for $k \in S$ to be a solution to the restricted
regression (defined below). We also show that $\hat{c}_k =
0$ satisfies the optimality condition of
optimization~\eqref{opt:alternate_opt_concave} for all $k \in S^c$.

\begin{definition}
\label{def:restricted_regression}
We define the \emph{restricted regression} problem 
\[
\min_{d_k} \frac{1}{n} \Big\| Y - \sum_{k \in S} \bar{\Delta}_k d_k \Big\|_2^2 + 
   \lambda_n \sum_{k \in S} \| \bar{\Delta}_k d_k \|_\infty \quad \trm{such that} \, d_{k,1}, \ldots, d_{k,n-1} \geq 0
\]
where we restrict the indices $k$ in
optimization \eqref{opt:alternate_opt} to lie in the set $S$ of true
relevant variables.
\end{definition}

\begin{theorem}[Deterministic setting]
\label{thm:deterministic}
Let $\{\hat{d}_k \}_{k \in S}$ be a minimizer of the restricted regression as defined above.
Let $\hat{r} \coloneqq Y - \sum_{k \in S} \bar{\Delta}_k \hat{d}_k$ be the restricted regression residual. 
Suppose for all $k\in S^c$, for all $i=1,\ldots,n$, $\lambda_n > | \frac{1}{2n}
\hat{r}^\tran \mathbf{1}_{(i:n)}|$ where $\mathbf{1}_{(i:n)}$ is 1 on
the coordinates of the 
$i$-th largest to the $n$-th largest entries of $X_k$ and 0
elsewhere.  Then the following two statements hold.
\begin{enumerate}
\item Let $\hat{d}_k = 0$ for $k \in S^c$.  Then
  \{$\hat{d}_k\}_{k=1,\ldots,p}$ is an optimal solution to
  optimization~\eqref{opt:alternate_opt}. Furthermore, any solution to
  the optimization program \eqref{opt:alternate_opt} must be zero on
  $S^c$.
\item For all $k \in S^c$, the solution $\hat{c}_k$ to optimization~\eqref{opt:alternate_opt_concave} must be zero.
\end{enumerate}

@@

\begin{enumerate}
\item The data $X^{(1)},\ldots, X^{(n)} \sim F$ are iid from
a distribution $F$ that is supported and strictly positive on $\mathcal{X}=[-b,b]^p$. 
\item The response is $Y = f_0(X) + W$ where $W$ is
  independent, zero-mean noise; thus $Y^{(i)} = f_0(X^{(i)}) + W^{(i)}$.
\item The regression function $f_0$ satisfies
$f_0(X) = f_0(X_{S_0})$, where $S_0 = \{1,\ldots,s_0\}$ is the set of
relevant variables.
\end{enumerate}

Let $\mathcal{C}^1$ denote the set of univariate convex functions
supported on $[-b,b]$, 
and let  $\mathcal{C}_1^{p}$ denote the set of convex additive functions
$\mathcal{C}_1^p \equiv \{ f \,:\, f = \sum_{k=1}^p f_k, \,
   f_k \in \mathcal{C}^1 \} $.  
Let $f^*(x) = \sum_{k=1}^p f^*_k(x_k)$ be the population risk
minimizer in $\mathcal{C}_1^p$, 
\begin{equation}
f^* \equiv \arg\min_{f \in \mathcal{C}_1^p} \E\big(f_0(X) - f^*(X)
\big)^2.
\end{equation}
Similarly, we define $\mh \mathcal{C}^1$ as the set of univariate concave functions supported on $[-b, b]$ and define
\begin{equation}
g^*_k = \arg\min_{g_k \in \mh \mathcal{C}^1} \E \big( f_0(X) - f^*(X)
- g_k(X_k) \big)^2.
\end{equation}
We let $S = \{ k = 1,\ldots,p \,:\, f^*_k \neq 0 \trm{ or } g^*_k \neq 0\}$. By additive faithfulness (Theorem~\ref{thm:acdc_faithful}), it must be that $S_0 \subset S$.

Each of our theorems will use a subset of the following assumptions:
\begin{packed_enum}
\item[A1:] $X_S, X_{S^c}$ are independent. 
\item[A2:] $f_0$ is convex and twice-differentiable. 
\item[A3:] $\|f_0\|_\infty \leq sB$ and $\| f^*_k \|_\infty \leq B$ for all $k$.
\item[A4:] $W$ is mean-zero sub-Gaussian, independent of $X$, with scale $\sigma$; i.e., for all $t \in \R$, $\E e^{t \epsilon} \leq e^{\sigma^2 t^2 / 2}$.
\end{packed_enum}
By assumption A1, $f^*_k$ is must be zero for $k\notin S$.
We define $\alpha_f, \alpha_g$ as a measure of the signal strength of the weakest variable:
\begin{align*}
\alpha_f &= \inf_{f \in \mathcal{C}_1^p \,:\, \exists k ,\, f^*_k \neq 0 \,\wedge\, f_k = 0} 
       \Big\{ \mathbb{E} \big( f_0(X) - f(X) \big)^2 - 
        \mathbb{E} \big( f_0(X) - f^*(X) \big)^2  \Big\}\\
\alpha_g &=   \min_{k \in S \,:\, g^*_k \neq 0}
      \Big\{ \mathbb{E} \big( f_0(X) - f^*(X) \big)^2 - 
    \mathbb{E} \big( f_0(X) - f^*(X) - g^*_k(X_k) \big)^2 \Big\}
\end{align*}
These play the role of the absolute value of the smallest nonzero
coefficient in the true linear model in lasso theory.  Intuitively, if
$\alpha_f$ is small, then it is easier to make a false omission in the
additive convex stage of the procedure. If $\alpha_g$ is small, then
it is easier to make a false omission in the decoupled concave stage
of the procedure.

@@

\begin{theorem}[Controlling false positives]
\label{thm:false_positive}
Suppose assumptions A1-A4 hold. Define $\tilde{\sigma} \equiv \max(\sigma, B)$,
and suppose 
\begin{equation}
\lambda_n \geq 8 s_0 \tilde{\sigma}  \sqrt{ \frac{\log^2 np}{n}}.
\end{equation}  
Then with probability at least $ 1 - \frac{12}{np}$, for all $k \in
S^c$, and for all $i'=1,\ldots, n$,
\begin{equation}
\Big| \frac{1}{2n}\hat{r}^\tran \mathbf{1}_{(i':n)_k} \Big| < \lambda_n
\end{equation}
and for all $k \in S^c$, both the AC solution $\hat{f}_k$, from optimization~\eqref{opt:alternate_opt}, and the DC solution $\hat{g}_k$, from optimization~\eqref{opt:alternate_opt_concave} are zero. 
\end{theorem}

The proof of Theorem~\ref{thm:false_positive} exploits independence of
$\hat{r}$ and $X_k$ from A1; when $\hat{r}$ and $X_k$ are independent,
$\hat{r}^\tran \mathbf{1}_{(i':n)}$ is the sum of $n - i' +1$ random
coordinates of $\hat{r}$.  We can then use concentration of
measure results for sampling without replacement to argue that $|
\frac{1}{n} \hat{r}^\tran\mathbf{1}_{(i':n)}|$ is small with high
probability. The full proof of Theorem~\ref{thm:false_positive} is in
Section~\ref{sec:false_positive_proof} of the Appendix.
